---
title: "Parsing PhyloAcc - Genomes by Origin"
author: "G. Bravo"
date: "02/16/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

<style type="text/css">

body{ /* Normal  */
      font-size: 14px;
  }
td {  /* Table  */
  font-size: 10px;
}
h1.title {
  font-size: 22px;
  color: DarkBlue;
  font-family: Arial, Helvetica, sans-serif;
}
h1 { /* Header 1 */
  font-size: 18px;
  font-weight: bold;
  font-family: Arial, Helvetica, sans-serif;
}
h2 { /* Header 2 */
  font-size: 16px;
  font-family: Arial, Helvetica, sans-serif;
  font-style: oblique;
}
h3 { /* Header 3 */
  font-size: 14px;
  font-family: Arial, Helvetica, sans-serif;
  font-style: oblique;
}
code.r{ /* Code block */
    font-size: 14px;
    font-family: Arial, Helvetica, sans-serif;
    font-stretch: expanded;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
    font-family: Arial, Helvetica, sans-serif;
}
</style>

## This is a script to parse results from a PhyloAcc run. It uses parsed and copied files by 05_parse_PhyloAcc.sh

# 1. Setting working directory and loading packages:

```{r message=FALSE, warning=FALSE}
setwd("/n/holyscratch01/edwards_lab/gbravo/Genomes/06_5_phyloacc/parsed_results")
library(tidyverse, lib.loc = "/usr/local/lib/R/site-library")
library(ggthemes)
library(GGMM)
library(seqinr)
library(gridExtra)
```

# 2. Source PhyloAcc functions. These were created by PhyloAcc developers and can be retrieved from their github:

```{r init, include= FALSE}
source("drawAlign_function.R")
```

# 3. Read in the tree and associated files necessary to run functions in drawAlign_function.R:

```{r treeData, cache=F}
#antbirds_all_corrected.mod and batch000_species_names.txt copied from input_data and outputs beforehand.
#scientific_names.txt was created in advance via nano or copied beforehand from a previous run.
treeData <- prepare_data(tree_path = "antbirds_all_corrected.mod", species_name = "batch000_species_names.txt", common_name = "scientific_names.txt")
treeData$tip <- treeData$tree$tip.label
```

# 4. Read in the likelihood and posterior probability information.

```{r, message=FALSE, warning = FALSE}
likantbirds_complete <- read_tsv("genome_location_combined_elem_lik.txt", col_names = c("key", "cnee", "loglik.null", "loglik.target", "loglik.full", "logBF1", "logBF2", "loglik_Max_M0", "loglik_Max_M1", "loglik_Max_M2")) %>%
  arrange(loglik.full) %>%
  distinct(cnee, .keep_all = T) %>%
  mutate(bf1 = as.numeric(loglik.target) - as.numeric(loglik.null), bf2 = as.numeric(loglik.target) - as.numeric(loglik.full)) %>%
  mutate(key = as.numeric(key))

zpostantbirds_complete <- read_tsv("genome_location_combined_postZ_M2.txt", col_types = cols(.default = "d"))
postaccantbirds_complete <- zpostantbirds_complete %>% rename(key = No.) # Renaming column so both tables can talk to each other using "key."

```

# 5. Plotting the distribution of acceleration (n_rate) and conservation (c_rate).

```{r, message=FALSE, warning = FALSE}
hist(postaccantbirds_complete$n_rate, breaks =100)
hist(postaccantbirds_complete$c_rate)
```


# 6. Filtering out elements with conservation rate greater than 0.6 (c_rate > 0.6)

```{r, message=FALSE, warning = FALSE}
cneeantbirds_complete <- inner_join(postaccantbirds_complete, likantbirds_complete, by = c("key" = "key")) %>%
  filter(c_rate <= 0.60)
```

# 7. Adding CNEE names from bed file info & reading the alignment

```{r, message=FALSE, warning = FALSE}
#galGal_final_merged_CNEEs_named.bed comes from folder of CNEE alignments
cneeBED <- read_tsv("galGal6_final_merged_CNEEs_named.bed", col_names = c("chr", "start", "end", "cnee"))
cneeantbirds_complete <- inner_join(cneeantbirds_complete, cneeBED, by = c("cnee" = "cnee"))
```

# 8. Labeling CNEEs that are accelerated according to different values of log-BF1 and log-BF2

```{r, message=FALSE, warning = FALSE}
# Evidence for acceleration: log-BF1 > 10 & log-BF2 > 0
# Overly conservative evidence for acceleration: log-BF1 > 20 & log-BF2 > 0

cnees_labeled_10_0 <- cneeantbirds_complete %>% mutate(Accel = bf1 > 10, Spec = bf1 > 10 & bf2 > 0) 
cnees_labeled_20_0 <- cneeantbirds_complete %>% mutate(Accel = bf1 > 20, Spec = bf1 > 20 & bf2 > 0)

table(cnees_labeled_10_0$Spec)["TRUE"]
table(cnees_labeled_20_0$Spec)["TRUE"]
```


# 9. Selecting Top Accelerating CNEEs

```{r, message=FALSE, warning = FALSE}
bf.1_10_0 <- cnees_labeled_10_0 %>%
  filter(Spec == T) %>%
  arrange(chr,start)
write.table(bf.1_10_0, "bf.1_10_0.txt", sep="\t", quote = FALSE, row.names = FALSE)

topZ_CNEE <- bf.1_10_0 %>% select(chr,start,end,cnee)
write.table(topZ_CNEE, "topZ_CNEEs.txt", sep="\t", quote = FALSE, row.names = FALSE)
```

